### 2021-04-10 assertions/doc/thesis-bernhardt/chapters/2-use_case_analysis.tex
Cost analysis + counterexamples

\section{Cost analysis}
\draft{Formalize the costs of checking a formula}
\begin{itemize}
\item Describe a general cost model for execution of a SC
\item Calculate costs for checking a formula locally (in a loop)
\item Calculate costs for checking a single element
\item Compare results -> how high is the ``boilerplate'' cost?
\item When do the costs of distributed checking drop below local checking?
\end{itemize}

>>>>

\section{Cost analysis}\label{sec:usecase_cost}
In order to be able to evaluate whether distributed assertion checking can indeed lower the cost in comparison to a local verification, the cost of checking a formula on the blockchain needs to be formalized. Since this thesis describes an implementation of the approach on the Tezos blockchain, the following costs refer to that specific gas model, s.t. the results of the cost analyses are comparable (Tezos and its gas model are described in more detail in \secref{sec:tezos}). \\
According to the gas model, the total gas consumption of a transaction $T(src,\, dst=C,\, arg, \, ...)$ to an originated contract $C(storage, \, parameter\, type, \, code)$ is calculated by \eqref{eq:gas_transaction}, where $I$ is the set of internal operations:
\begin{align}\label{eq:gas_transaction}
\begin{split}
GasConsumed(T) = \quad &base \, amount \\
+& \, gasRead(code) \\
+& \, gasRead(storage) \\
+& \, gasDeserialize(storage, \, arg) \\
+& \, gasParseType(parameter \, type, \, storage) \\
+& \, gasParseData(storage, \, arg) \\
+& \, gasParseCode(code) \\
+& \, gasInterpret(code, \, arg, \, storage) \\
+& \, gasUnparse(storage) \\
+& \, gasSerialize(storage) \\
+& \, gasWrite(storage) \\
+& \, \sum_{i \in I} GasConsumed(i)
\end{split}
\end{align}
Assuming that the storage is never increased, the amount of gas returned by the cost functions independent of the argument, i.e., the parameter value, remain constant for each $T$. The cost functions for the deserialization, parsing and interpretation, however, are functions of $arg$ that grow strictly monotonic with increasing sizes of $arg$. \\
Consider the following simple contracts --- they do not do anything useful, but can be used to approximate the amplification of costs with regard to $arg$. Contract \texttt{Foo} takes an integer, whose byte sequence does not increase significantly with higher values, thus the increase in overall cost is mainly accounted for by interpretation cost. Contract \texttt{Bar}, on the other hand, executes similar operations but expects a list of integers, which causes noticeably higher deserialization and parsing costs with increasing input size. 
\begin{lstlisting}[numbers=none, language=Solidity, caption=Simple dummy contract executing a loop]
contract Foo {
  function foo (int a) public {
    for (uint i = 0; i < a; i++) {
      assert(a == a)
	}
	// ...
  }
}
\end{lstlisting}
\begin{lstlisting}[numbers=none, language=Solidity, caption=Simple dummy contract iterating over a list]
contract Bar {
  function bar (int [] b) public {
    for (uint i = 0; i < b.length; i++) {
      assert(b[i] = b[i])
	}
	// ...
  }
}
\end{lstlisting}

\figref{fig:use_case_cost} shows the reported gas consumption of transactions calling the dummy contracts on the Delphinet test network running the protocol 007 Delphi \cite{tezos_docs}. As to be expected, the gradient of the cost function in the case of \texttt{Bar} is higher due to additional serialization and parsing costs, namely $\frac{dx}{dy} = 16$ compared to $\frac{dx}{dy} = 6$ for \texttt{Foo}.
\begin{figure}[t]
\centering
\includegraphics[width=0.9\linewidth]{figures/2-use_cases/cost_analysis}
\caption{Gas consumption of transactions calling \texttt{Foo} and \texttt{Bar} for increasing input sizes}
\label{fig:use_case_cost}
\end{figure}

A repeated cost analysis is carried out in \secref{sec:cost_analysis_distributed} in order to assess whether, or how, distributed assertion checking, as described in the following chapters, can achieve a reduction of these costs.

### 2021-04-10 assertions/doc/thesis-bernhardt/chapters/3-offline_design.tex
Cost analysis + counterexamples

\subsection{Alternatives to random testing}
For applications where higher reliability is crucial, an alternative to checking elements randomly could be to coordinate a systematic iteration of the domain. The implementation of such a coordination is not trivial though; validators can't simply be passed an individual value as input. Instead, possible approaches could be to implement a central instance that allocates an element in the search space to each validator, or use some unique and intrinsic attribute, like an identification number, as input for a mapping. The discussion if this approach would provide any advantages over a local validation of the property is left open \todo{Depends on cost analysis?}.

>>>>

\subsection{Alternatives to random testing}\label{sec:alt_random}
For applications where higher reliability is crucial, an alternative to checking elements randomly could be to coordinate a systematic iteration of the domain. The implementation of such a coordination is not trivial though; validators can't simply be passed an individual value as input. Instead, possible approaches could be to implement a central instance that allocates an element in the search space to each validator, or use some unique and intrinsic attribute, like an identification number, as input for a mapping. The discussion if this approach would provide any advantages over a local validation of the property is left open \todo{Depends on cost analysis?}.

### 2021-04-10 assertions/doc/thesis-bernhardt/chapters/4-offline_design_tezos.tex
Cost analysis + counterexamples

The function \texttt{generate\_loop} generates code that does the following:
\begin{enumerate}
\item Retrieve the assertion contract and parent contract addresses from storage
\item Calculate $t$
\item Build a list of $t$ internal operations that call the assertion contract
\item Return the list of emitted operations and the unchanged storage
\end{enumerate}
By passing the next id $i$ to the function, the access to the storage can be hard-wired into the generated code. The address of the parent contract is assumed to always be stored with the id 0. \secref{sec:prob_threshold} stated a formula for calculating $t_total$ with the parameters $n$, the size of the search space, and the probability threshold $c$. While $n$ can be calculated from the ranges of the random generator(s), $c$ must be given as parameter to the compiler through the CLI. After $t_total$ has been computed, a divison by 32, i.e., the number of validators, returns the upper bound of the loop. \\
\texttt{generate\_forward} can be generated from a simple template with the id as the only variant and only consists of  an access to the storage and the emission of an internal operation.

@@

\section{Cost analysis}
Tezos specific costs analysis of chosen orchestration strategy -> PROBLEM: high n might reach operation gas limit?

\section{Open questions}

\subsection{Representing counterexamples}
\draft{}
- the random number somehow has to be extracted from the contract execution, s.t. other validators can verify the counterexample
- failwith exposes top element of the stack in error message -> can be used to "return" the random values? Question for protocol design how to process it further
- if this is not possible, Michelson may have to be extended with a special failure message that can be handled differently
- assertion execution -> random number is generated
- if counterexample is found, the random number somehow has to be extracted from the contract execution, s.t. other validators can verify the counterexample
- open question how the validators can check the counterexample -> send piece of code with random values as parameter?
- incentives \& penalties part of the block chain protocol

>>>>

The function \texttt{generate\_loop} generates code that does the following:
\begin{enumerate}
\item Retrieve the assertion contract and parent contract addresses from storage
\item Calculate $t$
\item Build a list of $t$ internal operations that call the assertion contract
\item Return the list of emitted operations and the unchanged storage
\end{enumerate}
By passing the next id $i$ to the function, the access to the storage can be hard-wired into the generated code. The address of the parent contract is assumed to always be stored with the id 0. \secref{sec:prob_threshold} stated a formula for calculating $t_{total}$ with the parameters $n$, the size of the search space, and the probability threshold $c$. While $n$ can be calculated from the ranges of the random generator(s), $c$ must be given as parameter to the compiler through the CLI. After $t_{total}$ has been computed, the upper bound of the loop is calculated according to \eqref{eq:t_validator}. 
\begin{equation}\label{eq:t_validator}
t = \lceil t_{total} / \#validators \rceil
\end{equation}

\texttt{generate\_forward} can be generated from a simple template with the id as the only variant and only consists of  an access to the storage and the emission of an internal operation.

@@

\section{Cost analysis}\label{sec:cost_analysis_distributed}
What remains to be evaluated is the cost of checking a formula with the proposed approach, and how they relate to the costs for a local validation which were analysed in \secref{sec:usecase_cost}. The previous analysis defined \eqref{eq:gas_transaction} to calculate the gas consumption of a transaction, including the emitted operations. For each validator, the transaction calling the manager contract emits $t+1$ internal operations, whose gas consumptions are also dependent of the input parameter. Based on the $t+1$ additional base, as well as deserialization and parsing costs of the parameter alone, it can be concluded that the total gas consumption of $n$ total distributed test runs is higher than the cost of checking $n$ values locally. \\
The differential can be measured by implementing the orchestration scheme for the two dummy contracts given in \secref{sec:usecase_cost}. Appendix \ref{apx} contains the code of all three components for each of them. In order to measure the gas consumption for a comparable amount of executions as during the initial cost analysis, the probability threshold $c$ is set to $\frac{1}{e}$, s.t. that the necessary number of total test runs is $n$. However, it needs to be pointed out, that the actual number of test runs executed can only be a multiple of the number of validators (cf. \eqref{eq:t_validator}). \figref{fig:cost_distributed} shows the reported gas consumption per validator, i.e., a $32^{th}$ of the total cost for checking an assertion for the given threshold. Due to \eqref{eq:t_validator}, the cost function approximates a step function.
\begin{figure}[t]
\centering
\includegraphics[width=0.9\linewidth]{figures/4-offline_tezos/cost_analysis}
\caption{Gas consumption per validator of transactions calling \texttt{Foo} and \texttt{Bar} for increasing input sizes}
\label{fig:cost_distributed}
\end{figure}
The results show that the cost incurred by a single validator checking $t$ values is already considerably higher than by checking the property locally. Moreover, the cost-performance ratio is even worse, as the confidence that the property is indeed satisfied is only around 63\% with the given parameters. In conclusion to these results, it may be worth looking into some alternative approaches. \secref{sec:alt_random} already mentions the possibility to coordinate the validators for a systematic iteration of the search space. Alternatively, the proposed orchestration could be modified, s.t. the manager and assertion code are not implemented as separate contracts, but with one single contract. Instead of $(t+1)$ internal operations, one only has to pay for one internal operation that calls the parent contract, hence avoiding the base, deserialization and parsing costs for each test run. Similar to the monolithic orchestration approach, this increases the complexity of the compilation process.\\ Another idea worth exploring is the adaption of the blockchain protocol to avoid user costs for assertion checking, similar to the regular validation of blocks \cite{tezos_docs}. Instead of being paid by the user in gas, validators could be given a reward for checking an assertion, or a particularly high one for finding a counterexample.

\section{Representing counterexamples}
There still remains an open question concerning the proposed implementation that could be answered in detail within the scope of this thesis - the representation and validation of a counterexamples. The interaction scheme in \figref{fig:interaction_modular} suggests, that validators publish an instance of a counterexample to the network if they found one. However, it has yet to be specified how the counterexample is represented and double-checked by the other validators. With the \texttt{failwith} instruction, the top element of the stack can be exposed \cite{tezos_docs}, providing a way to return an interpretable expression from the interpretation context \cite{tezos_repo}. Thus, as a first approximation, the counterexample could be represented as a tuple of assertion code and a list of randomly generated values. For instance, the counterexample for the assumption ``p is a prime'' (cf. \eqref{eq:prime}) is represented by \\
\texttt{(n >= 2 \&\& n <= sqrt(p) \&\& p\%n = 0, <random value for n>)}. \\
For validation of the counterexample, this tuple, together with the respective operation hash of the initial transaction this counterexample refers to, is published to the network. The other validators then interpret the given code with the respective value(s) and either confirm or object the counterexample. As Tezos transmits data as byte sequences, the counterexample needs to be unparsed and serialized, or deserialized and parsed respectively. If the cost for checking assertions is paid for be the user (in form of gas), then these costs have to be added to the total costs as well.\\
In order to prevent denial-of-service attacks by spamming the network with false counterexamples, the protocol should implement penalties for such cases. Details about this and possible rewards for valid counterexamples should be considered in the protocol design.

### 2021-04-10 assertions/doc/thesis-bernhardt/chapters/appendix.tex
Cost analysis + counterexamples

\appendix
\chapter{Assertion Grammar in EBNF}\label{apx:grammar}
The following two versions of the grammar implement the prefix and infix notations for the assertion syntax. 
\todo{Rename mutez}

@@

>>>>

\appendix
\chapter{Michelson implementations of \texttt{Foo} and \texttt{Bar}}\label{apx:cost_analysis_contract}
The following contracts are the Michelson equivalent of the dummy contracts given in \secref{sec:usecase_cost}.
\section{Foo}
\lstinputlisting[basicstyle=\linespread{1.0}\fontfamily{lmr}\selectfont\small,
				 backgroundcolor=\color{cverbbg},
				 linewidth=14cm,
				 xleftmargin=0.5cm,
				 frame=lr,
				 framesep=8pt,
				 framerule=0pt,
				 captionpos=b,
				 numbers=none,
				 language=Michelson,
				 caption=Dummy Michelson contract executing a dynamic loop]{listings/foo.tz}

\section{Bar}
\lstinputlisting[basicstyle=\linespread{1.0}\fontfamily{lmr}\selectfont\small,
				 backgroundcolor=\color{cverbbg},
				 linewidth=14cm,
				 xleftmargin=0.5cm,
				 frame=lr,
				 framesep=8pt,
				 framerule=0pt,
				 captionpos=b,
				 numbers=none,
				 language=Michelson,
				 caption=Dummy Michelson contract iterating over a dynamic list]{listings/bar.tz}

\chapter{Assertion Grammar in EBNF}\label{apx:grammar}
The following two versions of the grammar implement the prefix and infix notations for the assertion syntax. 
\todo{Rename mutez}

@@

\chapter{Distributed Michelson implementation of \texttt{Foo} and \texttt{Bar}}\label{apx:foobar_distributed}
The following contracts are a restructuring of the contracts given in Appendix \ref{apx:cost_analysis_contract}, s.t. they implement the proposed orchestration strategy.

\section{\texttt{Foo}}
\lstinputlisting[basicstyle=\linespread{1.0}\fontfamily{lmr}\selectfont\small,
				 backgroundcolor=\color{cverbbg},
				 linewidth=14cm,
				 xleftmargin=0.5cm,
				 frame=lr,
				 framesep=8pt,
				 framerule=0pt,
				 captionpos=b,
				 numbers=none,
				 language=Michelson,
				 label=lst:foo_manager,
				 caption=Manager contract of \texttt{Foo}]{listings/foo_manager.tz}
\lstinputlisting[basicstyle=\linespread{1.0}\fontfamily{lmr}\selectfont\small,
				 backgroundcolor=\color{cverbbg},
				 linewidth=14cm,
				 xleftmargin=0.5cm,
				 frame=lr,
				 framesep=8pt,
				 framerule=0pt,
				 captionpos=b,
				 numbers=none,
				 language=Michelson,
				 label=lst:foo_manager,
				 caption=Assertion contract of \texttt{Foo}]{listings/foo_assertion.tz}
\lstinputlisting[basicstyle=\linespread{1.0}\fontfamily{lmr}\selectfont\small,
				 backgroundcolor=\color{cverbbg},
				 linewidth=14cm,
				 xleftmargin=0.5cm,
				 frame=lr,
				 framesep=8pt,
				 framerule=0pt,
				 captionpos=b,
				 numbers=none,
				 language=Michelson,
				 label=lst:foo_manager,
				 caption=Parent contract of \texttt{Foo}]{listings/foo_parent.tz}

\section{\texttt{Bar}}
\lstinputlisting[basicstyle=\linespread{1.0}\fontfamily{lmr}\selectfont\small,
				 backgroundcolor=\color{cverbbg},
				 linewidth=14cm,
				 xleftmargin=0.5cm,
				 frame=lr,
				 framesep=8pt,
				 framerule=0pt,
				 captionpos=b,
				 numbers=none,
				 language=Michelson,
				 label=lst:bar_manager,
				 caption=Manager contract of \texttt{Bar}]{listings/bar_manager.tz}
\lstinputlisting[basicstyle=\linespread{1.0}\fontfamily{lmr}\selectfont\small,
				 backgroundcolor=\color{cverbbg},
				 linewidth=14cm,
				 xleftmargin=0.5cm,
				 frame=lr,
				 framesep=8pt,
				 framerule=0pt,
				 captionpos=b,
				 numbers=none,
				 language=Michelson,
				 label=lst:bar_manager,
				 caption=Assertion contract of \texttt{Bar}]{listings/bar_assertion.tz}
\lstinputlisting[basicstyle=\linespread{1.0}\fontfamily{lmr}\selectfont\small,
				 backgroundcolor=\color{cverbbg},
				 linewidth=14cm,
				 xleftmargin=0.5cm,
				 frame=lr,
				 framesep=8pt,
				 framerule=0pt,
				 captionpos=b,
				 numbers=none,
				 language=Michelson,
				 label=lst:bar_manager,
				 caption=Parent contract of \texttt{Bar}]{listings/bar_parent.tz}
